---
title: "Generalized Linear Models"
author: "Flint"
date: 
output:
  html_document: default
  pdf_document: default
  word_document: default
---
----------------------------------------------------------------------------------------------
Abstract


Key Words: Treatment, Onychomycosis, terbinafine, itraconazole, Logistic regression
_________________________________________________________________________________________________



Group leader: Flint Chenjera
----------------------------------------------------------------------------------------------

Dependent variable: 

Independent variables: 


----------------------------------------------------------------------------------------------

House keeping
```{r warning=FALSE}
comment=NA # to remove all hashes
rm(list=ls())
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
library(nlme)
library(MASS)
library(tidyverse)
library(ggplot2)
library(nlme)
library(knitr)
library(kableExtra)
library(expss)
library(pander)
library(stargazer)
library(dplyr)
library(gmodels)
library(xtable)
library(survMisc)
library(TraMineR)
library(stargazer)
library(grid)
library(vcd)
library(epiDisplay)
opts_knit$set(root.dir = "C:/Users/user/Desktop/flint/MSc Statistics/GLM 503/Project")
```



#Load the data
```{r}
thedat <- read.table("toenail.dat",row.names=NULL)
 
dim(thedat)
#The original study consisted of 1200 observations and 6 variables
head(thedat)
#The original variables are not properly named and the values for categorical variables are not labeled
summary(thedat)
#from the looks of things, V2, V3 and V6 are binary variables.
```

#Data manupulation 
```{r}
#renaming variables
colnames(thedat) <- c("ID","healthclub","gender","month","length","trt")
head(thedat)
length(unique(thedat$id))

unique(thedat$healthclub) # 0 1
unique(thedat$gender)     # 1 0
unique(thedat$month)      # 0 1 2 3 6 12
unique(thedat$trt)        # 1 2   

#adding value labels 
thedat<-thedat %>%
  mutate(Treatment=ifelse(trt==1,"terbinane","itraconazole")) %>%
  mutate(Health_club=ifelse(healthclub==1,"more than once a week","once a week or less")) %>%
  mutate(Gender=ifelse(gender==1,"Male","Female")) %>%
  mutate(Month=ifelse(month==0, "Base month", 
                      ifelse(month==1,"1st month",
                             ifelse(month==2,"2nd month", 
                                    ifelse(month==3,"3rd month",
                                           ifelse(month==6,"6th month", "12th month"))))))

dat1<-                       
  filter(thedat, Month=="Base month"| Month=="12th month")
unique(dat1$Month)

boxpt<-boxplot(length ~ Treatment*month, data = thedat, ylab="length" , xlab="Treatment + month", col=c("gold","darkgreen"), names = c("it0","te0","it1","te1","it2","te2","it3","te3","it6","te6","it12","te12"))

data_wide <- reshape(dat1, idvar = "ID", timevar = "Month", direction = "wide")
summary(data_wide)

data_wide<-data_wide %>%
  mutate(differance=`length.12th month`- `length.Base month`) %>%
  mutate(success=ifelse(differance>=5,"Success","Failure"))

count1<-data_wide %>%
  count(success) %>%
mutate("%" = n / sum(n)*100)

count1  %>%   
  kable("html" , caption = 'Outcome Descriptives') %>%   
  kable_styling()
#kable(head(thedat), format = "markdown")
```



Is there evidence to suggest that these two oral medications differ in
their effects on toenail dermatophyte onychomycosis, as reflected by
the proportions of successfully treated patients at 12 months?
```{r}
#Graphical visualization
p<-ggplot(data=thedat, aes(x=month,y=length,group=ID ),alpha=0.90)+
geom_point(size=2,shape=18,color='navy blue') +
geom_line(size=1,color='navy blue') +
theme(strip.text.x = element_text(size =15, colour = "navy blue")) +
scale_x_continuous("Time (Month)",limits=c(0,12),breaks=thedat$month) +
scale_y_continuous("Length (mm)",limits=c(0,20))
  
p1<-p+ facet_wrap(~Treatment , nrow=1)
p2<-p+ facet_wrap(~Treatment+Gender , nrow=1)

#Analytical 
#table7<- data_wide %>%  table(`Treatment.12th month`, success) 

table1<- 
 table(data_wide$`Treatment.12th month`, data_wide$success) 

a1<-CrossTable(data_wide$`Treatment.12th month`, data_wide$success)
CrossTable(data_wide$`Treatment.12th month`, data_wide$`Gender.12th month`)
CrossTable(data_wide$`Treatment.12th month`, data_wide$`Health_club.12th month`)
```


##Regrssion models
```{r}
mod1=glm(factor(success)~1,family = binomial(link=logit),data=data_wide) 
#summary(mod1)
logistic.display(mod1)

mod2=glm(factor(success)~factor(`Treatment.12th month`),family = binomial(link=logit),data=data_wide)
#summary(mod2)
logistic.display(mod2)

#anova(mod1,mod2,test="LRT")

mod3=glm(factor(success)~factor(`Treatment.12th month`)+ `length.Base month`
        ,family = binomial(link=logit),data=data_wide)
#summary(mod3)
logistic.display(mod3)



#anova(mod2,mod3,test="LRT") 

mod4=glm(factor(success)~factor(`Treatment.12th month`)+ `length.Base month`+
           factor(`Health_club.12th month`) ,family = binomial(link=logit),data=data_wide)
#summary(mod4)
logistic.display(mod4)

#anova(mod3,mod4,test="LRT") ##---Likelihood ratio test between models

mod5=glm(factor(success)~factor(`Treatment.12th month`)+
      `length.Base month`+ factor(`Health_club.12th month`) + factor(`Gender.12th month`)   ,family = binomial(link=logit),data=data_wide)
#summary(mod5)
logistic.display(mod5) 

#anova(mod4,mod5,test="LRT") ##---Likelihood ratio test between models

#including an interaction term

#data_wide<-data_wide %>%
  #mutate(inter1= interaction(Treatment, Healthclub))

#mod4.1=glm(factor(success)~factor(`Treatment.12th month`)+ `length.Base month`+
#           factor(`Health_club.12th month`+ inter1) ,family = binomial(link=logit),data=data_wide)
#summary(mod4.1)


Treatment=ifelse(data_wide$`Treatment.12th month`=="terbinane",1,0)
Length=data_wide$`length.Base month`
Healthclub=ifelse(data_wide$`Health_club.12th month`=="once a week or less",1,0)
Gender=ifelse(data_wide$`Gender.Base month`=="Male",1,0)

coef4<-coef(mod4) ###--- Model parameter estimates

prm=coef(mod4)

predit.Func=function(prm,x1,x2,x3)
{
  pred1=exp(prm[1]+prm[2]*x1+prm[3]*x2+prm[4]*x3)/(1+exp(prm[1]+prm[2]*x1+prm[3]*x2+prm[4]*x3))
  pred1
}
pred1=predit.Func(prm=coef4,x1=Treatment, x2=Length, x3=Healthclub)

##three identical variable from model 4

data_wide<-data_wide %>%
  mutate(Indicator=ifelse(`Health_club.Base month`=="more than once a week" & `Treatment.Base month`=="terbinane", "Health Club:> 1 week & Treatment: terbinane", ifelse(`Health_club.Base month`=="more than once a week" & `Treatment.Base month`=="itraconazole", "Health Club:> 1 week & Treatment: itraconazole", ifelse(`Health_club.Base month`=="once a week or less" & `Treatment.Base month`=="itraconazole", "Health Club:<=1 week & Treatment: itraconazole", "Health Club:<=1 week & Treatment: terbinane")))) %>%
  mutate(pred1=pred1)

#Plot of predicted probabilities against different class of subjects

g1<-ggplot(data = data_wide, aes(x=`length.Base month`, y=pred1, group = Indicator, colour = Indicator, fill=NULL)) +
    geom_line(size = 1)+ 
  geom_point(size = 1.6) +
  scale_x_continuous("Lenght at Base month",limits=c(0, 10)) +
scale_y_continuous("Predicted Probability of success of a treatment",limits=c(0,1))
g1

##Combining the fine models produced above
stargazer(mod1, mod2, mod3, mod4, mod5 ,type="text", summary= TRUE, column.labels = c("Model", "Model", "Model", "Model", "Model"), covariate.labels = c("Treatment: terbinane", "Length at Base month", "Health club: once a week or less","Gender: Male", "Intercept"), style = "default",dep.var.labels.include=FALSE,dep.var.caption="Dependent variable: Treatment success")

####OR OR OR OR OR OR OR OR OR
```


## checking for change in health club visitations
```{r}
data_wide1 <- reshape(thedat, idvar = "ID", timevar = "Month", direction = "wide")
total1<-data_wide1
total1<-select(total1, healthclub0=`Health_club.Base month`, healthclub1=`Health_club.1st month`, healthclub2=`Health_club.2nd month`, healthclub3=`Health_club.3rd month`, healthclub6=`Health_club.6th month`, healthclub12=`Health_club.12th month`)


labels.total <- c("more than once a week", "more than once a week")
total1$"Wave 0" <- total1$healthclub0
total1$"Wave 1" <- total1$healthclub1
total1$"Wave 2" <- total1$healthclub2
total1$"Wave 3" <- total1$healthclub3
total1$"Wave 6" <- total1$healthclub6
total1$"Wave 12" <- total1$healthclub12


healthclub.seq <- seqdef(total1, c("Wave 0", "Wave 1","Wave 2","Wave 3","Wave 6", "Wave 12"), xtstep = 1, labels = labels.total)


fam.trate <- seqtrate(healthclub.seq)
round(fam.trate, 2)

```



